### 2019-08-29 manuscript/manuscript.tex
updated figures, ms text

\author{}

% \author[$\ast$,$\dag$,$1$]{Vince Buffalo}
% \author[$\dag$]{Graham Coop}
% \affil[$\ast$]{\footnotesize Population Biology Graduate Group}
% \affil[$\dag$]{\footnotesize Center for Population Biology, Department of Evolution and Ecology, University of California, Davis, CA 95616}
% \affil[$1$]{\footnotesize Email for correspondence: \href{mailto:vsbuffalo@ucdavis.edu}{vsbuffalo@ucdavis.edu}}

@@

We measure two types of covariances between allele frequency caused by
selection: temporal autocovariances, and across-replicate covariances. First,
positive temporal autocovariance in a neutral allele frequency's trajectory
occurs when the allele becomes associated with a high or low fitness
chromosomal background, and this association persists through the generations
due to linkage disequilibrium. As long as the direction of selection remains
constant, the fitness background is predictive of the direction in the neutral
allele's frequency changes, creating positive covariance. Even though these
magnitude of frequency changes at each site may be subtle, as they would be
under polygenic selection, cumulatively these perturb neighboring sites in a
predictable manner and build up temporal autocovariance which acts as a
genome-wide signal of linked selection. Second, if evolution occurs in
replicate populations undergoing convergent selection pressure, neutral sites
linked to fitness backgrounds shared across replicates are expected to change
in the same direction. This creates across-replicate covariance, which is a
measure of the extent to which convergent selection pressures across replicate
populations cause similar allele frequency changes. Finally, it is important to
note that under the null model where the fitness differences between
individuals are entirely random and non-heritable (e.g. when selection is not
acting), both forms of covariances are expected to be zero.

\begin{center}
  \begin{tabular}{l c c c c c c}
    Study & Species & Selection & Replicates & Pop. Size & Gens. & Timepoints \\ 
  \hline
    \textcite{Kelly2019-dc}  & \emph{D. simulans} & lab & 3 & $\sim$1100 & 14 & 2 \\ \hline
    \textcite{Barghi2019-qy} & \emph{D. simulans} & lab & 10 & $\sim$1000 & 60 & 10 \\ \hline
    \textcite{Castro2019-uk} & \emph{M. musculus} & \begin{tabular}{c}tibiae length\\control\end{tabular} & \begin{tabular}{c}2\\1\end{tabular} &\begin{tabular}{c}32\\28\end{tabular}  & 20 & 2 \\
 \hline
\end{tabular}
\caption{Table to test captions and labels}
\label{table:1}
\end{center}

We first analyzed \textcite{Barghi2019-qy}, an evolve-and-resequence study with
ten replicate populations exposed to a high temperature environment and evolved
for 60 generations, and sequenced every ten generations. Using the seven
timepoints and ten replicate populations, we estimated a bias-corrected $60
\times 60$ temporal-replicate variance-covariance matrix (see XXX for details).
Since each replicate population was sequenced every ten generations, the
timepoints $t_0 = 0$ generations, $t_1 = 10$ generations, $t_2 = 20$
generations, etc., lead to observed allele frequency changes across ten
generation blocks, $\Delta p_{t_0}, \Delta p_{t_1}, \ldots, \Delta p_{t_6}$.
Consequently, the ten temporal covariance matrices for each of the ten
replicate populations have off-diagonal elements of the form $\cov(\Delta
p_{t_0}, \Delta p_{t_1}) = \cov(p_{t_1} - p_{t_0}, p_{t_2} - p_{t_1}) =
\sum_{i=0}^{10} \sum_{j=10}^{20} \cov(\Delta p_i, \Delta p_j)$. Each diagonal
element has the form $\var(\Delta p_{t_0}) = \sum_{i=0}^{t_0} \var(\Delta
p_{i}) + \sum_{i \ne j}^{t_0} \cov(\Delta p_{i}, \Delta p_{j})$, and is thus a
combination of the effects of drift and selection, as both the variance in
allele frequency changes and cumulative temporal autocovariances terms increase
the variance in allele frequency. With sampling each generation, one could more
accurately partition the total variance in allele frequency change
\parencite{Buffalo2019-io}; while we cannot directly estimate the contribution
of linked selection to the variance in allele frequency change here, the
presence of a positive observed covariance between allele frequency change can
only be caused linked selection. Overall, we expect observed temporal
covariances between timepoints to be weak, some fraction of the additive
genetic variance and linkage disequilibrium that creates temporal
autocovariance decays in the ten generations between sequencing.

@@

\begin{align}
  \mathrm{cor}(\Delta p_s, \Delta p_t) = \frac{\E_{A\ne B} \left( \cov(\Delta p_{s,A}, \Delta p_{t,B}) \right)}{\E_{A\ne B} \left( \sqrt{\var(\Delta p_{s,A}) \var(\Delta p_{t,B})} \right)}
\end{align}
%
where $A$ and $B$ here are two replicate labels.

@@

Finally, we observed that in the longest run evolve-and-resequence study we
analyzed \parencite{Barghi2019-qy}, some temporal covariances appeared to
become negative at future timepoints (see the first two rows in Figure
\ref{fig:figure-1} A). This could be caused by advantageous fitness backgrounds
later becoming disadvantageous, either due to frequency-dependent selection,
recombination breaking up coupled beneficial and deleterious alleles, or an
environmental change in the direction of selection. However, the signal of a
change in the direction of selection localized to a particular region of the
genome may be washed out when we calculate genome-wide temporal covariances.
Averaging over such large numbers of loci is necessary to detect a signal of
covariance in allele frequency changes when both drift (inflated in
evolve-and-sequence studies which use small population sizes) and sampling
variance can create spurious covariances. To address this limitation, we
calculated temporal covariances over XXX Mb genomic windows, with on average
XXX loci per window.  While the covariance of each tile is noisy, we can
analyze the genome-wide distribution of tile covariances for changes through
time. We would like to compare the distribution of these tile covariances to a
neutral expectation, but a theoretically-determined neutral null distribution
would be inappropriate given complex dependencies across the genome due to LD
(which can occur over long physical distances in E&R and selection studies,
\cite{Nuzhdin2013-gf,Baldwin-Brown2014-cl}).

\begin{figure}[!ht]
  \centering
  \includegraphics[]{figures/figure-3.pdf}
  \caption{}
  \label{fig:figure-3}
\end{figure}

@@

@@

\section{Sampling Bias Corrections}

@@

\begin{align}
  \frac{d_0-1}{d_0} \frac{\var(\widetilde{p}_1 - \widetilde{p}_0)}{\widetilde{p}_0(1-\widetilde{p}_0)} - \frac{(d_0-1)}{d_0 (d_1 - 1)} \frac{\widetilde{p}_1(1-\widetilde{p}_1)}{\widetilde{p}_0(1-\widetilde{p}_0)} - \frac{1}{d_0} + \frac{C}{N}  &= \frac{\var(\Delta_{_H} p_0)}{p_0(1-p_0)} + \frac{1}{2N} 
\end{align}

\subsection{Individual and depth sampling process}

$X_t \sim \text{Binom}(n_t, p_t)$ where $X_t$ is the count of alleles and $n_t$
is the number of diploids sampled at time $t$. Then, these individuals are
sequenced at a depth of $d_t$, and $Y_t \sim \text{Binom}(d_t,
\nicefrac{X_t}{n_t})$ reads have the tracked allele. We let $\widetilde{p_t} =
\nicefrac{Y_t}{d_t}$ be the observed sample allele frequency. Then, the
sampling noise is

\begin{align}
  \var(\widetilde{p_t}|p_t) &= \E(\var(\widetilde{p_t} | X_t)) + \var(\E(\widetilde{p_t} | X_t)) \\
                            &= p_t(1-p_t) \left(\frac{1}{n_t} + \frac{1}{d_t} - \frac{1}{n_t d_t} \right).
\end{align}

(see also \cite{Jonas2016-ia}).

@@

\begin{align}
  \frac{n_t d_t}{(n_t-1) (d_t-1)} \widetilde{p_t}(1-\widetilde{p_t})
\end{align}

\begin{align}
  \var(\widetilde{p}_t - \widetilde{p}_0) &= 
  \frac{n_t d_t \widetilde{p}_t(1-\widetilde{p}_t)}{(n_t-1)(d_t-1)} \left(\frac{1}{n_t} + \frac{1}{d_t} - \frac{1}{n_t d_t} \right) + 
 \frac{n_0 d_0 \widetilde{p}_0(1-\widetilde{p}_0)}{(n_0-1)(d_0-1)} \left( \frac{1}{n_0} + \frac{1}{d_0} - \frac{1}{n_0 d_0}\right) + \\ & \nonumber\;\;\;\;\;\;
 \frac{n_0 d_0 \widetilde{p}_0(1-\widetilde{p}_0)}{(n_0-1)(d_0-1)}   \left[1 - \left(1-\frac{1}{2N}\right)^t \right]  - \frac{C}{N}  \frac{n_0 d_0 \widetilde{p}_0(1-\widetilde{p}_0)}{(n_0-1)(d_0-1)} + \\ \nonumber & \;\;\;\;\;\; \sum_{i=0}^{t-1} \var(\Delta_{_H} p_i)  + \sum_{0 \le i < j}^{t-1} \cov(\Delta p_i, \Delta p_j)  \\
                                                                                                                      &= \widetilde{p}_t(1-\widetilde{p}_t)\frac{d_t + n_t - 1}{(n_t-1)(d_t-1)} + 
 \widetilde{p}_0(1-\widetilde{p}_0)\frac{d_0 + n_0 - 1}{(n_0-1)(d_0-1)} + \\ & \nonumber\;\;\;\;\;\;
 \widetilde{p}_0(1-\widetilde{p}_0) \frac{n_0 d_0}{(n_0-1)(d_0-1)}  \left[1 - \left(1-\frac{1}{2N}\right)^t \right] - \frac{C}{N} \widetilde{p}_0(1-\widetilde{p}_0)\frac{n_0 d_0}{(n_0-1)(d_0-1)} 
 \\ \nonumber & \;\;\;\;\;\; + \sum_{i=0}^{t-1} \var(\Delta_{_H} p_i)  + \sum_{0 \le i < j}^{t-1} \cov(\Delta p_i, \Delta p_j) \
\end{align}

\subsection{Covariance Correction}

@@

With frequency collected at $T+1$ timepoints across $R$ replicate populations
at $L$ loci, we have $\mathbf{F}$ of allele frequencies, $\mathbf{D}$
multidimensional array of sequencing depths, and a $\mathbf{N}$
multidimensional array of the number of individuals sequenced, each of
dimension $R \times (T+1) \times L$.  We calculate the array $\mathbf{\Delta
F}$ which contains the allele frequency changes between adjacent generations,
and has dimension $R \times T \times L$.  The operation
$\flt(\mathbf{\Delta}\mathbf{F})$ flattens this array to a $(R \cdot T) \times
L$ matrix, such that rows are grouped by replicate, e.g. for timepoint $t$,
replicate $r$, and locus $l$ and allele frequencies $p_{t, r}$, for a single
locus the entries are

@@

\begin{align}
  \mathbf{B} = \frac{1}{L} \sum_{l=1}^L \frac{\mathbf{h}_l}{2} \circ \left( \frac{1}{\mathbf{d}_l} + \frac{1}{2\mathbf{n}_l} + \frac{1}{2\mathbf{d}_l \circ \mathbf{n}_l} \right)
\end{align}

where $\circ$ denotes elementwise product, and $\mathbf{h}_l$, $\mathbf{d}_l$,
and $\mathbf{n}_l$, are rows corresponding to locus $l$ of the unbiased
heterozygosity arrays $\mathbf{H}$, depth matrix $\mathbf{D}$, and number of
diploids matrix $\mathbf{N}$. The unbiased $R \times (T+1) \times L$
heterozygosity array can be calculated as

\begin{align}
  \mathbf{H} = \frac{2 \mathbf{D} \circ \mathbf{N} }{ (\mathbf{D}-1) \circ (\mathbf{N} -1)} \circ \mathbf{F} \circ (1-\mathbf{F})
\end{align}

where division here is elementwise. Thus, $\mathbf{B}$ is a $R \times (T+1)$
matrix. As explained in XXX, each temporal covariance submatrix
$\mathbf{Q}_{r,r}$ requires two corrections.

\begin{equation}
	\[
	Q_{i,j} =  
		\begin{dcases}
			Q_{t,s}' - b_{t} - b_{s}, & \text{if  } t = s \\
			Q_{t,s}' + b_{}, & \text{if  } |i - j| = 1 \\
		\end{cases}
	\]
\end{equation}

$Q_{i,j} = \cov(\Delta p_i, \Delta p_j)$

Additionally, in some study designs, the frequencies from the first timepoint

\begin{align}
  \mathbf{Q} = \mathbf{Q}' - \mathrm{diag}(\flt(\mathbf{B}_{\cdot,2:(T+1)})) - \mathrm{diag}(\flt(\mathbf{B}_{\cdot, 1:T})) + \mathrm{offdiag}_1(\flt(\mathbf{B}_{\cdot, 2:T})) + \mathrm{offdiag}_{-1}(\flt(\mathbf{B}_{\cdot, 2:T}))
\end{align}

where $\mathrm{diag}(\mathbf{x})$ is an operation that takes the vector
$\mathbf{x}$ and places them along the diagonal of a matrix. Similarly,
$\mathrm{offdiag}_{k}(\mathbf{x})$ places vector $\mathbf{x}$ along the
offdiagonal where for row $i$, and column $j$, $j-i = k$. We represent subset
of columns $s$ through $t$ of a matrix $\mathbf{B}$ as $\mathbf{B}_{\cdot,
s:t}$.

@@

\subsection{Correction}
\begin{figure}[!ht]
  \centering
  \includegraphics[]{figures/barghi-correction-plot.pdf}
  \caption{}
  \label{suppfig:correction}
\end{figure}

>>>>

\author[$\ast$,$\dag$,$1$]{Vince Buffalo}
\author[$\dag$]{Graham Coop}
\affil[$\ast$]{\footnotesize Population Biology Graduate Group}
\affil[$\dag$]{\footnotesize Center for Population Biology, Department of Evolution and Ecology, University of California, Davis, CA 95616}
\affil[$1$]{\footnotesize Email for correspondence: \href{mailto:vsbuffalo@ucdavis.edu}{vsbuffalo@ucdavis.edu}}

@@

% We measure two types of covariances between allele frequency caused by
% selection: temporal autocovariances, and across-replicate covariances. First,
% positive temporal autocovariance in a neutral allele frequency's trajectory
% occurs when the allele becomes associated with a high or low fitness
% chromosomal background, and this association persists through the generations
% due to linkage disequilibrium. As long as the direction of selection remains
% constant, the fitness background is predictive of the direction in the neutral
% allele's frequency changes, creating positive covariance. Even though these
% magnitude of frequency changes at each site may be subtle, as they would be
% under polygenic selection, cumulatively these perturb neighboring sites in a
% predictable manner and build up temporal autocovariance which acts as a
% genome-wide signal of linked selection. Second, if evolution occurs in
% replicate populations undergoing convergent selection pressure, neutral sites
% linked to fitness backgrounds shared across replicates are expected to change
% in the same direction. This creates across-replicate covariance, which is a
% measure of the extent to which convergent selection pressures across replicate
% populations cause similar allele frequency changes. Finally, it is important to
% note that under the null model where the fitness differences between
% individuals are entirely random and non-heritable (e.g. when selection is not
% acting), both forms of covariances are expected to be zero.

\begin{center}
  \begin{tabular}{l c c c c c c}
    Study & Species & Selection & Replicates & Pop. Size & Gens. & Timepoints \\ 
  \hline
    \textcite{Kelly2019-dc}  & \emph{D. simulans} & lab & 3 & $\sim$1100 & 14 & 2 \\ \hline
    \textcite{Barghi2019-qy} & \emph{D. simulans} & lab & 10 & $\sim$1000 & 60 & 10 \\ \hline
    \textcite{Castro2019-uk} & \emph{M. musculus} & \begin{tabular}{c}tibiae length\\control\end{tabular} & \begin{tabular}{c}2\\1\end{tabular} &\begin{tabular}{c}32\\28\end{tabular}  & 20 & 2 \\
 \hline
\end{tabular}
\end{center}

We first analyzed \textcite{Barghi2019-qy}, an evolve-and-resequence study with
ten replicate populations exposed to a high temperature lab environment and
evolved for 60 generations, and sequenced every ten generations. Using the
seven timepoints and ten replicate populations, we estimated a bias-corrected
genome-wide $6 \times 6$ temporal covariance matrix (a timepoint is lost in
calculating all adjacent allele frequency changes). This temporal covariance
matrix contains

$60 \times 60$ temporal-replicate variance-covariance matrix (see XXX for
details).

Since each replicate population was sequenced every ten generations,
the timepoints $t_0 = 0$ generations, $t_1 = 10$ generations, $t_2 = 20$
generations, etc., lead to observed allele frequency changes across ten
generation blocks, $\Delta p_{t_0}, \Delta p_{t_1}, \ldots, \Delta p_{t_6}$.
Consequently, the ten temporal covariance matrices for each of the ten
replicate populations have off-diagonal elements of the form $\cov(\Delta
p_{t_0}, \Delta p_{t_1}) = \cov(p_{t_1} - p_{t_0}, p_{t_2} - p_{t_1}) =
\sum_{i=0}^{10} \sum_{j=10}^{20} \cov(\Delta p_i, \Delta p_j)$. Each diagonal
element has the form $\var(\Delta p_{t_0}) = \sum_{i=0}^{t_0} \var(\Delta
p_{i}) + \sum_{i \ne j}^{t_0} \cov(\Delta p_{i}, \Delta p_{j})$, and is thus a
combination of the effects of drift and selection, as both the variance in
allele frequency changes and cumulative temporal autocovariances terms increase
the variance in allele frequency. With sampling each generation, one could more
accurately partition the total variance in allele frequency change
\parencite{Buffalo2019-io}; while we cannot directly estimate the contribution
of linked selection to the variance in allele frequency change here, the
presence of a positive observed covariance between allele frequency change can
only be caused linked selection. Overall, we expect observed temporal
covariances between timepoints to be weak, some fraction of the additive
genetic variance and linkage disequilibrium that creates temporal
autocovariance decays in the ten generations between sequencing.

@@

\begin{align}
  \label{eq:conv-corr}
  \mathrm{cor}(\Delta p_s, \Delta p_t) = \frac{\E_{A\ne B} \left( \cov(\Delta p_{s,A}, \Delta p_{t,B}) \right)}{\E_{A\ne B} \left( \sqrt{\var(\Delta p_{s,A}) \var(\Delta p_{t,B})} \right)}
\end{align}
%
where $A$ and $B$ here are two replicate labels.

@@

Finally, we observed that in the longest run evolve-and-resequence study we
analyzed \parencite{Barghi2019-qy}, some temporal covariances appeared to
become negative at future timepoints (see the first two rows in Figure
\ref{fig:figure-1} A). This could be caused by advantageous fitness backgrounds
later becoming disadvantageous, either due to frequency-dependent selection,
recombination breaking up coupled beneficial and deleterious alleles, or an
environmental change in the direction of selection. However, the signal of a
change in the direction of selection localized to a particular region of the
genome may be washed out when we calculate genome-wide temporal covariances.
Averaging over such large numbers of loci is necessary to detect a signal of
covariance in allele frequency changes when both drift (inflated in
evolve-and-sequence studies which use small population sizes) and sampling
variance can create spurious covariances. To address this limitation, we
calculated temporal covariances over 100kb genomic windows, with on average XXX
loci per window. While the covariance of each tile is noisy, we can analyze the
genome-wide distribution of tile covariances for changes through time. We would
like to compare the distribution of these tile covariances to a neutral
expectation, but using a theoretic neutral null distribution would be
inappropriate given complex dependencies across the genome due to linkage
disequilibria (which can occur over long physical distances in E\&R and
selection studies, \cite{Nuzhdin2013-gf,Baldwin-Brown2014-cl}).

Instead, we have developed a permutation-based procedure that constructs a null
distribution of the genome-wide covariances as they would look if only neutral
genetic drift was acting. We do this by randomly flipping the signs of the
allele frequency changes per-genomic window 1,000 times, which acts to destroy
the systematic covariances created by linked selection but creates a sampling
distribution of the covariances spuriously created by neutral genetic drift.
This empirical neutral null distribution is conservative in the sense that the
variances of the covariances are wider than expected under drift alone, since
drift also acts to inflate the magnitude of allele frequency change. By
flipping the sign at the block-level, we preserve the complex dependencies
between adjacent loci created by linkage disequilibrium; see Supplementary
Material Section \ref{supp:empirical-null} for more details on this approach.

Using the \textcite{Barghi2019-qy} study's data, we used the empirical neutral
null to investigate whether there was a general trend for the temporal
covariance $\cov(\Delta p_t, \Delta p_{t+k})$ to be positive for small $k$ (as
expected under selection), but then become systematically negative for larger
$k$ (as would occur under evolutionary processes that reverse the fitness of
regions across the genome). Such a change in the fitness direction of
haplotypes would not be discernible from genome-wide covariances like those in
Figure \ref{fig:figure-1}. We find (Figure \ref{fig:figure-3} A and B), pooling
across all replicates (see Supplementary Figure
\ref{suppfig:barghi-offset-replicate-panels} for individuals replicates), that
windowed temporal covariances between close timepoints are skewed positive (a
heavy right tail), while between more distant timepoints these windowed
temporal covariances tend to shift to become more negative (a heavy left tail).

\begin{figure}[!ht]
  \centering
  \includegraphics[]{figures/figure-3.pdf}

  \caption{A, B: The distribution of temporal covariances calculated in 100kb
    genomic windows from the \textcite{Barghi2019-qy} study, plotted alongside
    an empirical neutral null distribution created by recalculating the
    windowed covariances on sign permuting the allele frequency changes. In
    subfigure A, windowed covariances $\cov(\Delta p_t, \Delta p_{t+k})$ are
    separated by $k=2 \times 10$ generations and in subfigure B the covariances
    are separated by $k=4 \times 10$ generations; each $k$ is an off-diagonal
    from the variance diagonal of the temporal covariance matrix (see cartoon
    of upper-triangle of covariance matrix in subfigures A and B, where the red
    diagonal is the variance, and the dark gray indicates which off-diagonal of
    the covariance matrix is plotted in the histograms). C: The lower and upper
    tail probabilities of the observed windowed covariances, at 20\% and 80\%
    quintiles of the empirical neutral null distribution, for varying time
    between allele frequency changes (i.e. which off-diagonal $k$). The
    confidence intervals are  $95\%$ block-bootstrap confidence interval, and the
    light gray dashed line indicates the 20\% tail probability expected under the
    neutral null.}

    \label{fig:figure-3} 
\end{figure}

@@

\section{Acknowledgments}

Doc Edge.

@@

\section{Estimator Bias Correction}
\subsection{Correcting variance bias with a single depth sampling process}
\label{supp:depth-var-corr}

@@

\begin{align}
  \label{supp:eqn-depth-only-correction}
  \frac{d_0-1}{d_0} \frac{\var(\widetilde{p}_1 - \widetilde{p}_0)}{\widetilde{p}_0(1-\widetilde{p}_0)} - \frac{(d_0-1)}{d_0 (d_1 - 1)} \frac{\widetilde{p}_1(1-\widetilde{p}_1)}{\widetilde{p}_0(1-\widetilde{p}_0)} - \frac{1}{d_0} + \frac{C}{N}  &= \frac{\var(\Delta_{_H} p_0)}{p_0(1-p_0)} + \frac{1}{2N} 
\end{align}

\subsection{Correcting variance bias with individual and depth sampling processes}
\label{supp:ind-depth-var-corr}

Here, we extend the sampling bias correction described above to handle two
binomial sampling processes: one as individuals are binomially sampled from the
population, and another as reads are binomially sampled during sequencing.
(see also \cite{Jonas2016-ia}). Let $X_t \sim \text{Binom}(n_t, p_t)$ where
$X_t$ is the count of alleles and $n_t$ is the number of diploids sampled at
time $t$. Then, these individuals are sequenced at a depth of $d_t$, and $Y_t
\sim \text{Binom}(d_t, \nicefrac{X_t}{n_t})$ reads have the tracked allele. We
let $\widetilde{p_t} = \nicefrac{Y_t}{d_t}$ be the observed sample allele
frequency. Then, the sampling noise is

\begin{align}
  \var(\widetilde{p_t}|p_t) &= \E(\var(\widetilde{p_t} | X_t)) + \var(\E(\widetilde{p_t} | X_t)) \\
                            &= p_t(1-p_t) \left(\frac{1}{n_t} + \frac{1}{d_t} - \frac{1}{n_t d_t} \right)
\end{align}

@@

\begin{align}
  \frac{n_t d_t}{(n_t-1) (d_t-1)} \widetilde{p_t}(1-\widetilde{p_t}).
\end{align}

Replacing this unbiased heterozygosity into our expression above, the total
sample variance is

\begin{align}
  \var(\widetilde{p}_t - \widetilde{p}_0) &= 
  \frac{n_t d_t \widetilde{p}_t(1-\widetilde{p}_t)}{(n_t-1)(d_t-1)} \left(\frac{1}{n_t} + \frac{1}{d_t} - \frac{1}{n_t d_t} \right) + 
 \frac{n_0 d_0 \widetilde{p}_0(1-\widetilde{p}_0)}{(n_0-1)(d_0-1)} \left( \frac{1}{n_0} + \frac{1}{d_0} - \frac{1}{n_0 d_0}\right) + \\ & \nonumber\;\;\;\;\;\;
 \frac{n_0 d_0 \widetilde{p}_0(1-\widetilde{p}_0)}{(n_0-1)(d_0-1)}   \left[1 - \left(1-\frac{1}{2N}\right)^t \right]  - \frac{C}{N}  \frac{n_0 d_0 \widetilde{p}_0(1-\widetilde{p}_0)}{(n_0-1)(d_0-1)} + \\ \nonumber & \;\;\;\;\;\; \sum_{i=0}^{t-1} \var(\Delta_{_H} p_i)  + \sum_{0 \le i < j}^{t-1} \cov(\Delta p_i, \Delta p_j).  \\
                                                                                                                      % &= \widetilde{p}_t(1-\widetilde{p}_t)\frac{d_t + n_t - 1}{(n_t-1)(d_t-1)} + 
 % \widetilde{p}_0(1-\widetilde{p}_0)\frac{d_0 + n_0 - 1}{(n_0-1)(d_0-1)} + \\ & \nonumber\;\;\;\;\;\;
 % \widetilde{p}_0(1-\widetilde{p}_0) \frac{n_0 d_0}{(n_0-1)(d_0-1)}  \left[1 - \left(1-\frac{1}{2N}\right)^t \right] - \frac{C}{N} \widetilde{p}_0(1-\widetilde{p}_0)\frac{n_0 d_0}{(n_0-1)(d_0-1)} 
 % \\ \nonumber & \;\;\;\;\;\; + \sum_{i=0}^{t-1} \var(\Delta_{_H} p_i)  + \sum_{0 \le i < j}^{t-1} \cov(\Delta p_i, \Delta p_j). 
\end{align}

As with equation \eqref{supp:eqn-depth-only-correction}, we can rearrange this
to get a biased-correct estimate of the variance in allele frequency change
between adjacent generations, $\var(\Delta p_t)$.

\subsection{Covariance Correction}
\label{supp:cov-corr}

@@

Now, we extend the bias corrections for single locus variance and covariance
described in Supplementary Material Sections \ref{supp:depth-var-corr},
\ref{supp:ind-depth-var-corr}, and \ref{supp:cov-corr} to multiple sampled
loci. With frequency collected at $T+1$ timepoints across $R$ replicate
populations at $L$ loci, we have $\mathbf{F}$ of allele frequencies,
$\mathbf{D}$ multidimensional array of sequencing depths, and a $\mathbf{N}$
multidimensional array of the number of individuals sequenced, each of
dimension $R \times (T+1) \times L$.  We calculate the array $\mathbf{\Delta
F}$ which contains the allele frequency changes between adjacent generations,
and has dimension $R \times T \times L$.  The operation
$\flt(\mathbf{\Delta}\mathbf{F})$ flattens this array to a $(R \cdot T) \times
L$ matrix, such that rows are grouped by replicate, e.g. for timepoint $t$,
replicate $r$, and locus $l$ and allele frequencies $p_{t, r}$, for a single
locus the entries are

@@

\begin{align}
  \mathbf{B} = \frac{1}{L} \sum_{l=1}^L \frac{\mathbf{h}_l}{2} \circ \left( \frac{1}{\mathbf{d}_l} + \frac{1}{2\mathbf{n}_l} + \frac{1}{2\mathbf{d}_l \circ \mathbf{n}_l} \right)
\end{align}
%
where $\circ$ denotes elementwise product, and $\mathbf{h}_l$, $\mathbf{d}_l$,
and $\mathbf{n}_l$, are rows corresponding to locus $l$ of the unbiased
heterozygosity arrays $\mathbf{H}$, depth matrix $\mathbf{D}$, and number of
diploids matrix $\mathbf{N}$. The unbiased $R \times (T+1) \times L$
heterozygosity array can be calculated as

\begin{align}
  \mathbf{H} = \frac{2 \mathbf{D} \circ \mathbf{N} }{ (\mathbf{D}-1) \circ (\mathbf{N} -1)} \circ \mathbf{F} \circ (1-\mathbf{F})
\end{align}
%
where division here is elementwise. Thus, $\mathbf{B}$ is a $R \times (T+1)$
matrix. As explained in Supplementary Material Section
\ref{supp:ind-depth-var-corr} and \ref{supp:cov-corr}, the temporal variances
and covariances require bias corrections, meaning each temporal covariance
submatrix $\mathbf{Q}_{r,r}$ requires two corrections. For an element
$Q_{r,t,s} = \cov(\Delta p_t, \Delta p_s)$ of the temporal covariance submatrix
for replicate $r$, $\mathbf{Q}_{r,r}$, we apply the following correction

\begin{equation}
	Q_{r,t,s} =  
		\begin{dcases}
			Q_{r,t,s}' - b_{r,t} - b_{r,t+1}, & \text{if  } t = s \\
      Q_{r,t,s}' + b_{r,\max(t,s)}, & \text{if  } |t - s| = 1 \\
		\end{dcases}
\end{equation}

where $b_{r,t}$ is element in row $r$ and column $t$ of $\mathbf{B}$.
Additionally, in some study designs, a single timepoint is shared for the
initial generation across replicates. In this case, the sampling noise is
shared between

@@

To infer the uncertainty of covariance, convergence correlation, and $G(t)$
estimates, we used a block bootstrap procedure. This is a version of the
bootstrap that resamples blocks of data points, rather than individual data
points, to infer the uncertainty of an statistic in the presence of unknown
correlation structure between data. With genome-wide data, linkage
disequilibria between sites creates complex and unknown dependencies between
variants. The estimators used in this paper are predominantly ratios, e.g.
temporal-replicate covariance standardized by half the heterozygosity (XXX),
$G(t)$ which is the ratio of covariance to total variance (XXX), and the
convergence correlation (equation \eqref{eq:conv-corr}). In these cases, we can
exploit the linearity of the expectation to make the bootstrap procedure more
computationally efficient, by pre-calculating the statistics of the ratio's
numerator and denominator, $N(\mathbf{x}_i)$ and $D(\mathbf{x}_i)$, on the data
$\mathbf{x}_i$ for all blocks $i \in \{1, 2, \ldots, W\}$ in the genome. Then
we draw $W$ bootstrap samples with replacement, and compute the estimate for
bootstrap sample $b$ with an average weighted by the number of loci in all
sampled blocks,

\begin{align}
  \tilde{\theta}_b = \sum_{i=1}^W w_i \frac{N(\mathbf{x}_i)}{D(\mathbf{x}_i)}
\end{align}
%
Note that computing the ratio of averages rather than the average of a ratio is
a practice common for population genetic statistics like $F_{ST}$
\parencite{Bhatia2013-zy}. With these $B$ bootstrap estimates, we calculate the
$\nicefrac{\alpha}{2}$ and $1-\nicefrac{\alpha}{2}$ quantiles, which we use to
estimate the $1-\alpha = 95\%$ pivot confidence intervals (p. 33
\cite{Wasserman2006-jl}, p. 194 \cite{Davison2013-oy}) throughout the paper,

\begin{align}
  C_\alpha = \left(2 \widehat{\theta} - q_{1-\nicefrac{\alpha}{2}}, 2 \widehat{\theta} - q_{\nicefrac{\alpha}{2}} \right).
\end{align}
%
where $\widehat{\theta}$ is the estimate, and $q_x$ is bootstrap quantile for
probability $x$.

\subsection{The Empirical Neutral Null Windowed Covariance Distribution}
\label{supp:empirical-null}

\section{Supplementary Figures}

\subsection{Bias Correction for \textcite{Barghi2019-qy}}

We have investigated the effectiveness of our correction on real data by
exploiting the relationship between sampling depth and the magnitude of the
variance and covariance biases, and comparing the observed variances and
covariances before and after correction. We plot the variance and covariance
(between adjacent timepoints) before and after the bias correction against the
average sample depth in 100kb genomic windows in Figure
\ref{suppfig:barghi-correction}. Overall, we find the correction strongly

\begin{figure}[!ht]
  \centering
  \includegraphics[]{figures/barghi-correction-plot.pdf}

  \caption{The variance and covariances calculated in 100kb genomic windows
    plotted against average depth in a window before and after bias correction.
    Each panel has a least-squares estimate between the variance and
    covariance, and the average depth. Overall, the bias correction corrects
    sampling bias in both the variance and covariance such that the
    relationship with depth is constant. Colors indicate the different
    chromosomes of \emph{D. simulans}; we have excluded the X chromosome
    (yellow points) and chromosome 4 points (green points to far right) from
    the regression due to large differences in average coverage.}

  \label{suppfig:barghi-correction}
\end{figure}

\subsection{\textcite{Barghi2019-qy} Empirical Null and Windowed Covariance Distributions}

\begin{figure}[!ht]
  \centering
  \includegraphics[]{figures/barghi-offset-panels.pdf}

  \caption{The distribution of temporal covariances calculated across 100kb
    genomic windows from \textcite{Barghi2019-qy}'s study (orange) and the
    block sign permuted empirical neutral null distribution of the windowed
    covariances (blue). Each panel shows these windowed covariances and the
    empirical null distribution for covariances $\cov(\Delta p_t, \Delta p_{t+k})$,
  $k$ is the number of generations between allele frequency changes.}
  \label{suppfig:barghi-empnull-tilecovs}
\end{figure}

\begin{figure}[!ht]
  \centering
  \includegraphics[]{figures/barghi-offset-replicate-panels.pdf}

  \caption{The distribution of windowed temporal covariances alongside the
    empirical neutral null for five randomly sampled replicates (columns), for
    $k=2$ (first row) and $k=5$ (second row). The main figure of the paper
    pools all replicate window and empirical neutral null covariances; we show
    here the windowed temporal covariances tend to shift from being positive (a
    heavier right tail) to become more negative (a heavier left tail) through
    time within particular replicates.}

  \label{suppfig:barghi-offset-replicate-panels}
\end{figure}

\subsection{\textcite{Barghi2019-qy} Tail Probabilities for Windowed Covariances Distributions}

\begin{figure}[!ht]
  \centering
  \includegraphics[]{figures/barghi-tailprobs-panels.pdf}

  \caption{ }
  \label{suppfig:barghi-tailprobs-panels}
\end{figure}
